---
title: "Essay2"
output: html_document
---
```{r packages}
library(dplyr)
```

# Sysnopsis: describes and summarizes your analysis in at most 10 complete sentences.

Data from the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database were analised to determine the natural disasters which had the greatest consequences in the U.S. 
It has been revealed that Tornado's cause the greatest damage in total numbers, both to population health and economic consequences. 


## Question 1
Q1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

## Question 2
Q2. Across the United States, which types of events have the greatest economic consequences?

# Data Processing: how the data were loaded into R and processed for analysis.
```{r loading data, cache = TRUE}
setwd("/Users//shota/git//Coursera-Data-Science/Reproducible research/")
data <- read.csv("./data/repdata-data-StormData.csv")
fatalities <- with(data, tapply(FATALITIES, EVTYPE, sum))
injuries <- with(data, tapply(INJURIES, EVTYPE, sum))
```

```{r processing1}
fatalities <- sort(fatalities, decreasing = T)
injuries <- sort(injuries, decreasing = T)
```

```{r processing 2}
property <- with(data, tapply(PROPDMG, EVTYPE, sum))
crop <- with(data, tapply(CROPDMG, EVTYPE, sum))


total <- rep(NA, length(property))
names(total) <- names(property)
total <- property + crop
property <- sort(property, decreasing = T)
crop <- sort(crop, decreasing = T)
total <- sort(total, decreasing = T)
```

# Results
```{r question1}
head(fatalities)
head(injuries)
par(mfrow = c(2,1))
barplot(fatalities[1:3], ylab = "total fatalities")
barplot(injuries[1:3], ylab = "total injuries")
```

Toranado's are the most harmful in respect to population health.

```{r question2}
head(property)
head(crop)
head(total)
par(mfrow = c(1,1))
barplot(total[1:3], ylab = "total damage")
```

Tornado's have the greatest economic consequence, due to it's magnitude in propety damage
