---
title: "week2"
output: html_document
---

Markdown
### heading
# Heading
## Secondary Heading
### even smaller heading

### Unordered lists
- first
- second

### Ordered lists
1. First
3. Third
2. Second

Doesn't matter which order you lay them out

### Syntax
#### Links
[RStudio](http://www.rstudio.com/)

#### Advance Links  
I spend so much time reading [Simply Statistics][1]!  

  [1]: http://simplystatistics.org/ "Simply Statistics"

Here's an inline link to [Google](http://www.google.com/).
Here's a reference-style link to [Google][1].
Here's a very readable link to [Yahoo!][yahoo].

  [1]: http://www.google.com/
  [yahoo]: http://www.yahoo.com/

Here's a <span class="hi">[poorly-named link](http://www.google.com/ "Google")</span>.
Never write "[click here][^2]".
Visit [us][web].

  [^2]: http://www.w3.org/QA/Tips/noClickHere (Advice against the phrase "click here")    
   
  [web]: http://stackoverflow.com/ "Stack Overflow"
  
#### New lines
two spaces  create a new line

### What is markdown?  
- you can concentrate on the content  
- r markdown is markdown for r, we'll use knitr



```{r}
library(datasets)
summary(airquality)
```

Let's make a plot of the data

```{r, echo=TRUE}
plot(airquality)
```

Here's a regression model of ozone, wind, solar radiation, and temperature

```{r}
library(stats)
fit <- lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```

### Literate programming

knitr is good for most things, but not for long/specific documents

#### complicated way of knitting
```{r knitting, echo=TRUE, error=TRUE, eval=FALSE}
library(knitr)
setwd("/Users//shota/git//Coursera-Data-Science/Reproducible research/")
knit2html("week2.Rmd")
browseURL("week2.html")
```

the produced markdown document includes the result in the console

#### hiding results
```{r hidden, echo=FALSE, results='hide'}
set.seed(1)
x <- rnorm(100)
mean(x)
```


#### Inline text computations
```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`. My favorite random number is `r rand`.

#### Incorporating graphics
```{r graphics}
x <- rnorm(100)
y <- x + rnorm(100, sd=0.5)
```

Here's the scatter plot
```{r scatterplot, fig.height=4}
par(mar=c(5,4,2,1), las=1)
plot(x, y, main = "My simulated data")
```

### Making tables
```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```

Here is a table of regression coefficients
```{r showtable, results='asis'}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")
```

### Setting Global options
```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "hide")
```
First simulated data
```{r simulateddata, echo=TRUE}
x <- rnorm(100)
y <- x + rnorm(100, sd=0.5)
```

Here's the scatterplot

```{r scatterplot2, fig.height=4}
par(mar=c(5,4,2,1), las=1)
plot(x, y, main = "My simulated data")
```

### Caching computations
- when a computation takes a long time
- cache = TRUE on a chunk caches it
- results are loaded after the first run
