---
title: "week3"
output: html_document
---

```{r}
setwd("/Users//shota/git//Coursera-Data-Science/Reproducible research/")
```

### Hierarchy of information  
- Title
- Abstract
- Body/Results
- Materials
- Code

### Sending mails to a boss
- Shorter the better, push everything in the subject.  
- Ask yes/no
- Put keywords
- Add deadline

### Attachements  
- knitr
- github repo

## Reproducible Reseach Checklist  
### Start with good science
- garbage in garbage out
- simple questions
- work with good people

### Don't do anything by hand  
- editing excel
- moving data around your computer
- downloading data from the website (do it by code)

### Don't point and click(GUI)
- it's not reproducible
- good for exploratory analysis

### Teach a computer
- for download

```{r}
download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/00275/ Bike-Sharing-Dataset.zip", "ProjectData/Bike-Sharing-Dataset.zip")
```

### Keeping track of your Environment  

```{r}
sessionInfo()
```


