---
title: "week2"
output: html_document
---

Markdown
### heading
# Heading
## Secondary Heading
### even smaller heading

### Unordered lists
- first
- second

### Ordered lists
1. First
3. Third
2. Second

Doesn't matter which order you lay them out

### Syntax
#### Links
[RStudio](http://www.rstudio.com/)

#### Advance Links  
I spend so much time reading [Simply Statistics][1]!  

  [1]: http://simplystatistics.org/ "Simply Statistics"

Here's an inline link to [Google](http://www.google.com/).
Here's a reference-style link to [Google][1].
Here's a very readable link to [Yahoo!][yahoo].

  [1]: http://www.google.com/
  [yahoo]: http://www.yahoo.com/

Here's a <span class="hi">[poorly-named link](http://www.google.com/ "Google")</span>.
Never write "[click here][^2]".
Visit [us][web].

  [^2]: http://www.w3.org/QA/Tips/noClickHere (Advice against the phrase "click here")    
   
  [web]: http://stackoverflow.com/ "Stack Overflow"
  
#### New lines
two spaces  create a new line

### What is markdown?  
- you can concentrate on the content  
- r markdown is markdown for r, we'll use knitr



```{r}
library(datasets)
summary(airquality)
```

Let's make a plot of the data

```{r, echo=TRUE}
plot(airquality)
```

Here's a regression model of ozone, wind, solar radiation, and temperature

```{r}
library(stats)
fit <- lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```


