---
title: "Statistical Inference week3"
output: pdf_document
---

## T distruibution

$$ 
\frac{\overline{X} - \mu} {S / \sqrt{n}}
$$

- will be a t distribtion, not a regular distruibution  
- t always has a wider bit at the edges.  
- In large scale, it's nearly the same
- for poisson, other stuff are available

## Sleep data

```{r sleep}
data(sleep)
head(sleep)
library(ggplot2)
g <- ggplot(sleep, aes(x = group, y = extra, group = factor(ID)))
g <- g + geom_line(size = 1, aes(colour = ID)) + geom_point(size =10, pch = 21, fill = "salmon", alpha = .5)
g

g1 <- sleep$extra[1:10]; g2 <- sleep$extra[11:20]
difference <- g2 - g1
mn <- mean(difference); s <- sd(difference); n <- 10

mn + c(-1,1) * qt(.975, n-1)* s / sqrt(n)
t.test(difference)
t.test(g2, g1, paired = T)
t.test(extra ~ I(relevel(group,2)), paired = T, data = sleep)
```

Checking the 95% confidence interval for the mean of the paried test.  
The t test in 4 sentences.

## Independent group t confidence trials

$$
\overline{Y} - \overline{X} \pm t_{n_x + n_y -2, 1-\alpha/2} S_p (\frac{1}{n_x} + \frac{1}{n_y})^{1/2}
$$

This is what happens for 2 groups

The variance
$$
S^2_p = \{(n_x - 1)S^2_x + (n_y - 1)S^2_y \} / (n_x + n_y - 2)
$$

- this assumes constant variance among groups

calculate from a book  
8 oral contraceptive users, 21 controlls  
Xoc = 132.86, soc = 15.34
Xc = 127.44, sc = 18.23

```{r blood pressure}
sp <- sqrt((7 * 15.34^2 + 20*18.23^2) / (8 + 21 - 2))
132.86 - 127.44 + c(-1,1) * qt(.975, 27) * sp * (1/8 + 1/21)^.5
```

The answer is [-9.52, 20.4]  
This contains 0, you can't rule out that they're the same

#### Sleep patients again

```{r sleep as grouped}
n1 <- length(g1); n2 <- length(g2)
sp <- sqrt( ((n1 - 1) * sd(g1) ^2 + (n2 - 1) * sd(g2)^2) / (n1 + n2 - 2))
md <- mean(g2) - mean(g1)
semd <- sp * sqrt(1/n1 + 1/n2)
rbind(
        md + c(-1,1) * qt(.975, n1 + n2 -2) * semd,
        t.test(g2, g1, paired = F, var.equal = T)$conf,
        t.test(g2, g1, paired = T)$conf
)
```

The answer  
           [,1]     [,2]  
[1,] -0.2038740 3.363874  
[2,] -0.2038740 3.363874  
[3,]  0.7001142 2.459886  

In a paired test, if you take g2 - g1, the answer is the 3rd row.

```{r chickweight}
library(datasets); data(ChickWeight); library(reshape2)
## define weight gain or loss
wideCW <- dcast(ChickWeight, Diet + Chick ~ Time, value.var = "weight")
names(wideCW)[-(1:2)] <- paste("time", names(wideCW)[-(1:2)], sep = "")
library(dplyr)
wideCW <- mutate(wideCW, gain = time21 - time0)

```

### Plotting the raw data
```{r chickplot}
g <- ggplot(ChickWeight, aes(x = Time, y = weight, 
                             colour = Diet, group = Chick))
g <- g + geom_line()
g <- g + stat_summary(aes(group = 1), geom = "line", fun.y = mean, size = 1, col = "black")
g <- g + facet_grid(. ~ Diet)
g
```

The 4 the diet looks faster. Let's check that.

```{r gain by diet}
g <- ggplot(wideCW, aes(x = factor(Diet), y = gain, fill = factor(Diet)))
g <- g + geom_violin(col = "black", size = 2)
g
```

A t interval

```{r chick t}
wideCW14 <- subset(wideCW, Diet %in% c(1,4))
rbind(
        t.test(gain ~ Diet, paired = F, var.equal = T, data = wideCW14)$conf,
        t.test(gain ~ Diet, paried = F, var.equal = F, data = wideCW14)$conf
)
```

#### Unequal Variances

$$
\overline{Y} - \overline{X} \pm t_{df} * (\frac{s^2_x}{n_x} + \frac{s^2_y}{n_y}) ^ {1/2}
$$

t.test(var.equal=F) does the job


