---
title: "Statistical Inference week1"
output: html_document
---

#### Probability density functions
This is the probability density for a triangle, we use as a sample.

```{r probability density functions}
pbeta(c(0.4, 0.5, 0.6), 2, 1)
```

#### Quantiles  
- 95% percent quantile: about 95% of the density lies below that point
- q stands for quantile

```{r quantiles}
qbeta(0.5, 2, 1)
```

## Finding the center of mass

```{r mass}
library(UsingR); data(galton); library(ggplot2); library(reshape2)
longGalton <- melt(galton, measure.vars = c("child", "parent"))
g <- ggplot(longGalton, aes(x = value)) + geom_histogram(aes(y = ..density..,  fill = variable), binwidth=1, colour = "black") + geom_density(size = 2)
g <- g + facet_grid(. ~ variable)
g
```

## Using Manipulate

```{r manipulate}
library(manipulate)
library(ggplot2)
myHist <- function(mu){
    g <- ggplot(galton, aes(x = child))
    g <- g + geom_histogram(fill = "salmon", 
      binwidth=1, aes(y = ..density..), colour = "black")
    g <- g + geom_density(size = 2)
    g <- g + geom_vline(xintercept = mu, size = 2)
    mse <- round(mean((galton$child - mu)^2), 3)  
    g <- g + labs(title = paste('mu = ', mu, ' MSE = ', mse))
    g
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))
```

The MSE is lowest at the center of mass

## Unbiased

If the distribtution is inbiased, the mean of the average of random samples will be centered at the original distruibution.

```{r test}
x <- 1:4
p <- x/sum(x)
temp <- rbind(x, p)
rownames(temp) <- c("X", "Prob")
temp
sum(x*p)

0.3 * 0.75 / (0.3 * 0.75 + 0.7 * 0.52)
```

# Week 2

```{r confidence intervals}
library(UsingR)
data(father.son)
x <- father.son$sheight
(mean(x) + c(-1, 1) * qnorm(0.975) * sd(x)/sqrt(length(x)))/12
```

```{r binom}
binom.test(56, 100)$conf.int
```


```{r simulation}
n <- 20
pvals <- seq(0.1, 0.9, by=0.05)
nosim <- 1000
coverage <- sapply(pvals, function(p){
        phats <- rbinom(nosim, prob = p, size = n)/n
        ll <- phats - qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        ul <- phats + qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        mean(ll < p & ul > p)
})
```

The plot isn't so good

```{r better simulation}
n <- 100
pvals <- seq(0.1, 0.9, by=0.05)
nosim <- 1000
coverage <- sapply(pvals, function(p){
        phats <- rbinom(nosim, prob = p, size = n)/n
        ll <- phats - qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        ul <- phats + qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        mean(ll < p & ul > p)
})
```

```{r change phat}
n <- 20
pvals <- seq(0.1, 0.9, by=0.05)
nosim <- 1000
coverage <- sapply(pvals, function(p){
        phats <- (rbinom(nosim, prob = p, size = n) + 2)/(n + 4)
        ll <- phats - qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        ul <- phats + qnorm(0.975) * sqrt(phats * (1 - phats)/n)
        mean(ll < p & ul > p)
})
```