---
title: "Regression Models week 2"
output: pdf_document
---

# The linear model
$$
Y_i = \beta_1 X_{1i} + \beta_2 X_{2i} + \dots + \beta_p X_{pi} + \epsilon_i = 
\sum\limits_{k=1}^n X_{ik}\beta_j + \epsilon_i
$$

- basically remove beta1 from beta2, beta2 from beta1 

# Result
$$
\hat \beta_1 = \frac{\sum_{i=1}^n e_{i, Y | X_2} e_{i, X_1 | X_2}}{\sum_{i=1}^n e_{i, X_1 | X_2}^2}
$$

# Exapmle
$Y_{i} = \beta_1 X_{1i} + \beta_2 X_{2i}$ where $X_{2i} = 1$ is an intercept term.

```{r example}
n = 100;
x = rnorm(n)
x2 = rnorm(n)
x3 = rnorm(n)
y = 1 + x + x2 + x3 + rnorm(n, sd = .1)
ey = resid(lm(y ~ x2 + x3))
ex = resid(lm(x ~ x2 + x3))
sum(ey*ex) / sum(ex^2)
coef(lm(ey ~ ex - 1))
coef(lm(y~ x + x2 + x3))
```

# Quiz
Consider the following data with x as the predictor and y as as the outcome.

x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
Give a P-value for the two sided hypothesis test of whether β1 from a linear regression model is 0 or not.
```{r q1}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
fit <- lm(y ~ x)
fit
summary(fit)$coefficients
coef(fit)
n <- length(x)
sigma <- sum(resid(lm(y~x))^2)/(n-2)
pt(coef(fit)[2]/sqrt(sigma), n-2, lower.tail = F)
plot(x,y)
abline(fit)
pt(0.7224/0.223, n-2, lower.tail = F) * 2
resid(fit)
resid(fit)^2
sqrt(sum(resid(fit)^2)/7)
```

# Q3
In the mtcars data set, fit a linear regression model of weight (predictor) on mpg (outcome). Get a 95% confidence interval for the expected mpg at the average weight. What is the lower endpoint?
```{r q3}
data(mtcars)
fit <- with(mtcars, lm(mpg ~ wt))
with(mtcars, plot(mpg ~ wt))
summary(fit)$coefficients
fit
weight <- mean(mtcars$wt)
n <- 32
sigma <- sum(resid(fit)^2)/(n-2)
sd <- sqrt(sigma)
sigma_b <- sd/var(mtcars$wt)
37.285 + weight*(-5.344 + c(-1,1) * qt(.975, n-2) * sigma_b)
37.285 + weight*-5.344

attach(mtcars)
fit <- lm(mpg ~ wt)
newdata = data.frame(wt=weight)
predict(fit, newdata, interval = "confidence")
detach(mtcars)
```

# Q5
Consider again the mtcars data set and a linear regression model with mpg as predicted by weight (1,000 lbs). A new car is coming weighing 3000 pounds. Construct a 95% prediction interval for its mpg. What is the upper endpoint?
```{r q5}
w <- 3000/1000
?predict
predict(fit,w)
predict(fit)
a <- 24.89
attach(mtcars)
newdata = data.frame(wt = 3)
predict(fit, newdata, interval = "prediction")
```

# Q6
Consider again the mtcars data set and a linear regression model with mpg as predicted by weight (in 1,000 lbs). A “short” ton is defined as 2,000 lbs. Construct a 95% confidence interval for the expected change in mpg per 1 short ton increase in weight. Give the lower endpoint.
```{r q6}
n <- 32
sigma <- sum(resid(fit)^2)/(n-2)
sd <- sqrt(sigma)
summary(fit)$coefficients
var(wt)
sigma_b <- sd/var(wt)
2*(-5.344 + c(-1,1) * qt(.975, df = fit$df) * sigma_b)
```

# Q9
Refer back to the mtcars data set with mpg as an outcome and weight (wt) as the predictor. About what is the ratio of the the sum of the squared errors, ∑ni=1(Yi−Y^i)2 when comparing a model with just an intercept (denominator) to the model with the intercept and slope (numerator)?
```{r q9}
fit1 <- lm(mpg ~ 1)
fit1
e1 <- sum(resid(fit1)^2)
fit
e2 <- sum(resid(fit)^2)
e2/e1
```
