---
title: "Regression Models week3"
output:
  pdf_document: default
---

# Swiss fertility data
```{r swiss}
library(datasets)
data(swiss)
require(stats)
require(graphics)
pairs(swiss, panel = panel.smooth, main = "Swiss data", col = 3 + (swiss$Catholic>50))
?swiss
summary(lm(Fertility ~ . , data = swiss)) #. is all the variables
```

- agriculture is -0.1721
```{r agriculture}
summary(lm(Fertility ~ Agriculture, data = swiss))$coefficients
```

Adjustment reversed the effect.

Let's simulate this
```{r simulate}
n <- 100
x2 <- 1:n 
x1 <- .1 * x2 + runif(n -.1, .1); 
y = - x1 + x2 + rnorm(n, sd = .01)
summary(lm(y ~ x1))$coef
summary(lm(y ~ x1 + x2))$coef
```

- This can happen.
- which is the correct model?

# Include unnecessary stuff
```{r unneccssary}
z <- swiss$Agriculture + swiss$Education
lm(Fertility ~ . + z, data = swiss)
```

- z can't add any information
- z becomes NA

# Dummy variables

- it depends how you choose your binary variable
* $Y_i = \beta_0 + X_{i1} \beta_1 + X_{i2} \beta_2 + \epsilon_i$.

# Insect sprays
```{r, echo = FALSE, fig.height=5, fig.width=5}
require(datasets);data(InsectSprays); require(stats); require(ggplot2)
g = ggplot(data = InsectSprays, aes(y = count, x = spray, fill  = spray))
g = g + geom_violin(colour = "black", size = 2)
g = g + xlab("Type of spray") + ylab("Insect count")
g

```

```{r insect code}
summary(lm(count~spray, data = InsectSprays))$coef
summary(lm(count ~ 
             I(1 * (spray == 'B')) + I(1 * (spray == 'C')) + 
             I(1 * (spray == 'D')) + I(1 * (spray == 'E')) +
             I(1 * (spray == 'F'))
           , data = InsectSprays))$coef
```

# what if you include all six
```{r all six}
summary(lm(count ~ 
   I(1 * (spray == 'B')) + I(1 * (spray == 'C')) +  
   I(1 * (spray == 'D')) + I(1 * (spray == 'E')) +
   I(1 * (spray == 'F')) + I(1 * (spray == 'A')), data = InsectSprays))$coef
```

## What if we omit the intercept?
```{r, echo= TRUE}
summary(lm(count ~ spray - 1, data = InsectSprays))$coef
library(dplyr)
summarise(group_by(InsectSprays, spray), mn = mean(count))
```

## Reordering the levels
```{r}
spray2 <- relevel(InsectSprays$spray, "C")
summary(lm(count ~ spray2, data = InsectSprays))$coef
```
---

- You can also do it manualy if you try hard enough

# Hunger
```{r hunger}
hunger <- read.csv("/Users/shotashimizu/git/courses/07_RegressionModels/02_02_multivariateExamples/hunger.csv")
hunger <- hunger[hunger$Sex != "Both sexes",]
head(hunger)
lm1 <- lm(hunger$Numeric ~ hunger$Year)
plot(hunger$Year, hunger$Numeric, pch =19, col = "blue")
lines(hunger$Year, lm1$fitted, lwd = 1, col = "darkgrey")

plot(hunger$Year, hunger$Numeric, pch =19)
points(hunger$Year, hunger$Numeric, pch = 19, col = ((hunger$Sex == "Male")*1 + 1))
```

# Remember the linear model
$$H v_i = b_0 + b_1 Y_i + e_i$$

# Color by male/female

```{r male femae}
lmM <- lm(hunger$Numeric[hunger$Sex == "Male"] ~ hunger$Year[hunger$Sex == "Male"])
lmF <- lm(hunger$Numeric[hunger$Sex == "Female"] ~ hunger$Year[hunger$Sex == "Female"])
lines(hunger$Year[hunger$Sex == "Male"], lmM$fitted, col = "black", lwd = 3)
lines(hunger$Year[hunger$Sex == "Female"], lmF$fitted, col = "red", lwd = 3)
```

- assume that there is a differnce in the intercept between male/female

```{r lm both}
lmBoth <- lm(hunger$Numeric ~ hunger$Year + hunger$Sex)
plot(hunger$Year, hunger$Numeric, pch = 19)
points(hunger$Year, hunger$Numeric, pch = 19, col = ((hunger$Sex == "Male")*1 + 1))
abline(c(lmBoth$coeff[1], lmBoth$coeff[2]), col = "red", lwd = 3)
abline(c(lmBoth$coeff[1] + lmBoth$coeff[3], lmBoth$coeff[2]), col = "black", lwd = 3)
```

- this means intercept is gender but not the slope

# two lines, differnet slopes
```{r slopes}
lmBoth <- lm(hunger$Numeric ~ hunger$Year + hunger$Sex + hunger$Sex * hunger$Year)
plot(hunger$Year, hunger$Numeric, pch = 19)
points(hunger$Year, hunger$Numeric, pch = 19, col = ((hunger$Sex == "Male")*1 + 1))
abline(c(lmBoth$coeff[1], lmBoth$coeff[2]), col = "red", lwd = 3)
abline(c(lmBoth$coeff[1] + lmBoth$coeff[3], lmBoth$coeff[2] + lmBoth$coeff[4]), col = "black", lwd = 3)
summary(lmBoth)
```

# Income and year
```{r income year}
```

# Regression example
```{r aha}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(runif(n/2), runif(n/2)); 
beta0 <- 0; beta1 <- 2; tau <- 1; sigma <- .2
y <- beta0 + x * beta1 + t * tau + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```

- Things are clear here, the diffence in the two lines are the treatment effect

# Simulation 2

```{r, fig.height=5, fig.width=5, echo = FALSE, results='hide'}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(runif(n/2), 1.5 + runif(n/2)); 
beta0 <- 0; beta1 <- 2; tau <- 0; sigma <- .2
y <- beta0 + x * beta1 + t * tau + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```

- no overlap
- no difference in the intercept
- not comparable

## Simulation 3
```{r, fig.height=5, fig.width=5, echo = FALSE, results='hide'}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(runif(n/2), .9 + runif(n/2)); 
beta0 <- 0; beta1 <- 2; tau <- -1; sigma <- .2
y <- beta0 + x * beta1 + t * tau + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```

- the importance of adjustment

### Some things to note in this simulation
* Marginal association has red group higher than blue.
* Adjusted relationship has blue group higher than red.
* Group status related to X.
* There is some direct evidence for comparing red and blue
holding X fixed.

## Simulation 4
```{r, fig.height=5, fig.width=5, echo = FALSE, results='hide'}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(.5 + runif(n/2), runif(n/2)); 
beta0 <- 0; beta1 <- 2; tau <- 1; sigma <- .2
y <- beta0 + x * beta1 + t * tau + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```

---
## Discussion
### Some things to note in this simulation
* No marginal association between group status and Y.
* Strong adjusted relationship.
* Group status not related to X.
* There is lots of direct evidence for comparing red and blue
holding X fixed.

## Simulation 5
```{r, fig.height=5, fig.width=5, echo = FALSE, results='hide'}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(runif(n/2, -1, 1), runif(n/2, -1, 1)); 
beta0 <- 0; beta1 <- 2; tau <- 0; tau1 <- -4; sigma <- .2
y <- beta0 + x * beta1 + t * tau + t * x * tau1 + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t + I(x * t))
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2] + coef(fit)[4], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```

---
## Discussion
### Some things to note from this simulation
* There is no such thing as a group effect here. 
  * The impact of group reverses itself depending on X.
  * Both intercept and slope depends on group.
* Group status and X unrelated.
  * There's lots of information about group effects holding X fixed.
  
# Swiss
- data for several beta's
```{r swiss again}
data(swiss)
par(mfrow = c(2,2))
fit <- lm(Fertility ~ . , data = swiss)
plot(fit)
