---
title: "Regression Models week3"
output:
  pdf_document:
    latex_engine: lualatex
---

# Swiss fertility data
```{r swiss}
library(datasets)
data(swiss)
require(stats)
require(graphics)
pairs(swiss, panel = panel.smooth, main = "Swiss data", col = 3 + (swiss$Catholic>50))
?swiss
summary(lm(Fertility ~ . , data = swiss)) #. is all the variables
```

- agriculture is -0.1721
```{r agriculture}
summary(lm(Fertility ~ Agriculture, data = swiss))$coefficients
```

Adjustment reversed the effect.

Let's simulate this
```{r simulate}
n <- 100
x2 <- 1:n 
x1 <- .1 * x2 + runif(n -.1, .1); 
y = - x1 + x2 + rnorm(n, sd = .01)
summary(lm(y ~ x1))$coef
summary(lm(y ~ x1 + x2))$coef
```

- This can happen.
- which is the correct model?

# Include unnecessary stuff
```{r unneccssary}
z <- swiss$Agriculture + swiss$Education
lm(Fertility ~ . + z, data = swiss)
```

- z can't add any information
- z becomes NA

# Dummy variables

- it depends how you choose your binary variable
* $Y_i = \beta_0 + X_{i1} \beta_1 + X_{i2} \beta_2 + \epsilon_i$.

# Insect sprays
```{r, echo = FALSE, fig.height=5, fig.width=5}
require(datasets);data(InsectSprays); require(stats); require(ggplot2)
g = ggplot(data = InsectSprays, aes(y = count, x = spray, fill  = spray))
g = g + geom_violin(colour = "black", size = 2)
g = g + xlab("Type of spray") + ylab("Insect count")
g

```

```{r insect code}
summary(lm(count~spray, data = InsectSprays))$coef
summary(lm(count ~ 
             I(1 * (spray == 'B')) + I(1 * (spray == 'C')) + 
             I(1 * (spray == 'D')) + I(1 * (spray == 'E')) +
             I(1 * (spray == 'F'))
           , data = InsectSprays))$coef
```

# what if you include all six
```{r all six}
summary(lm(count ~ 
   I(1 * (spray == 'B')) + I(1 * (spray == 'C')) +  
   I(1 * (spray == 'D')) + I(1 * (spray == 'E')) +
   I(1 * (spray == 'F')) + I(1 * (spray == 'A')), data = InsectSprays))$coef
```

## What if we omit the intercept?
```{r, echo= TRUE}
summary(lm(count ~ spray - 1, data = InsectSprays))$coef
library(dplyr)
summarise(group_by(InsectSprays, spray), mn = mean(count))
```

## Reordering the levels
```{r}
spray2 <- relevel(InsectSprays$spray, "C")
summary(lm(count ~ spray2, data = InsectSprays))$coef
```
---

- You can also do it manualy if you try hard enough

# Hunger
```{r hunger}
hunger <- read.csv("/Users/shotashimizu/git/courses/07_RegressionModels/02_02_multivariateExamples/hunger.csv")
hunger <- hunger[hunger$Sex != "Both sexes",]
head(hunger)
lm <- lm(hunger$Numeric ~ hunger$Year)
plot(hunger$Year, hunger$Numeric, pch =19, col = "blue")
```

# Remember the linear model
$$H v_i = b_0 + b_1 Y_i + e_i$$

# Residuals
- what are outliers?
- some outliers have more force. if it's without other x's, it'll have more force.

# Tools
- ?influence.measures
- use knowing the context
- the 10,10 is creating a lot of noise

# Graphs of residuals
- dfbetas and dfhats show the magnitude of influence of a single plot
- the orly owl shows the importance of residuals
- the swiss data,,,

# Selecting models (Multiple Variables)
- variance inflation, add more to the model and variance goes up
- if the variable is correlated, more variance inflation
- temp <- apply(betas, 1, var); temp[2 : 3] / temp[1]
- this is inflation

# 
```{r variance inflation}
n <- 100; nosim <- 1000;
x1 <- rnorm(n)
x2 <- rnorm(n)
x3<- rnorm(n)
betas <- sapply(1:nosim, function(i){
  y <- x1 + rnorm(n, sd = .3)
  c(coef(lm(y ~ x1))[2],
    coef(lm(y ~ x1 + x2))[2],
    coef(lm(y ~ x1 + x2 + x3))[2])
})
round(apply(betas, 1, sd), 5)
```

