---
title: "Practical Machine learning"
output: html_document
---

```{r input}
setwd("git/Coursera-Data-Science/Practical Machine Learning/")
df <- read.csv("pml-training.csv")

inBuild <- createDataPartition(y=df$classe,p=0.7,list=F)
validation <- df[-inBuild,]
buildData <- df[inBuild,]

inTrain <- createDataPartition(y=buildData$classe,p=0.7,list=F)
training <- buildData[inTrain,]
testing <- buildData[-inTrain,]

final <- read.csv("pml-testing.csv")
str(training)
str(testing)
names(training)
table(training$classe)
```

# fit 

## a simple model will be used
- probably better for the computation

```{r simple}
which(colnames(training) == "gyros_belt_x")
which(colnames(training) == "magnet_belt_z")
which(colnames(training) == "gyros_arm_x")
which(colnames(training) == "magnet_arm_z")
which(colnames(training) == "classe")
subtraining <- training[,c(37:45,60:68,160)]

fitModTree <- train(classe ~.,data=subtraining,method="rpart")
pred <- predict(fitModTree,newdata=testing)
answer3 <- pred
answer3
```

## find out how to cross validate


## fit a model by random forest, using the predictors
- roll_belt -> - skewness_roll_belt
- min_yaw_belt
- gyros_belt_x -> total_accel_arm
- gyros_arm_x -> skewness_yaw_arm  
- roll_dumbell -> skewness_pitch_dumbell

## selct the predictors in data frame

```{r all but NA}
which(colnames(training) == "roll_belt")
which(colnames(training) == "skewness_roll_belt")
which(colnames(training) == "min_yaw_belt")
which(colnames(training) == "gyros_belt_x")
which(colnames(training) == "total_accel_arm")
which(colnames(training) == "gyros_arm_x")
which(colnames(training) == "skewness_yaw_arm")
which(colnames(training) == "roll_dumbbell")
which(colnames(training) == "skewness_pitch_dumbbell")

subtraining <- training[,c(8:15,23,37:49,60:74,84:91,160)]

fitModNA <- train(classe ~.,data = subtraining,method="rpart")
pred <- predict(fitModNA, newdata=testing)
table(pred)
fitModNA$finalModel
pred
answer5 <- pred
```

## select meaningful
roll_belt
yaw_belt
gyros_belt_x
yaw_arm
gyros_arm_x
magnet_arm_z

```{r with roll}
which(colnames(training) == "roll_belt")
which(colnames(training) == "yaw_belt")
which(colnames(training) == "gyros_belt_x")
which(colnames(training) == "yaw_arm")
which(colnames(training) == "gyros_arm_x")
which(colnames(training) == "magnet_arm_z")
which(colnames(training) == "gyros_belt_x")
which(colnames(training) == "magnet_belt_z")
which(colnames(training) == "gyros_arm_x")
which(colnames(training) == "magnet_arm_z")
which(colnames(training) == "classe")
subtraining <- training[,c(8:10,3,37:48,60:68,160)]

library(ggplot2)
library(caret)
library(dplyr)
fitMod <- train(classe ~.,data = subtraining,method="rpart")
pred <- predict(fitMod, newdata=testing)
table(pred)
fitMod$finalModel
pred
answer1 <- pred
```
answer1
 [1] A A B C A E B B A A B B B A E A A B D A
 forest with 26 predictors
 
# random forest wth 19 predictors

```{r rf}
fitMod <- train(classe ~ ., data=subtraining,method="rf")
pred <- predict(fitMod,newdata=testing)
pred
answer2 <- pred
answer2
```

> answer2
 [1] B A B A A E D B A A B C B A E E A B B B
 random forest with 19predictors
 
# Model with 19 predictors only

gyros_belt_x
magnet_belt_z
gyros_arm_x
magnet_arm_z
classe

```{r select meaningful tree 19}
which(colnames(training) == "gyros_belt_x")
which(colnames(training) == "magnet_belt_z")
which(colnames(training) == "gyros_arm_x")
which(colnames(training) == "magnet_arm_z")
which(colnames(training) == "classe")
subtraining <- training[,c(37:45,60:68,160)]

fitModTree <- train(classe ~.,data=subtraining,method="rpart")
pred <- predict(fitModTree,newdata=testing)
answer3 <- pred
answer3
```
 answer3
 [1] A A A A A D D D A A A D D A D A A D A A
 forest with 19 predictors

# Full tree model

```{r full}
fitModFull <- train(classe ~.,data=training,method="rpart")
pred <- predict(fitModFull, newdata=testing)
answer4 <- pred
answer4
fitModFull$finalModel
```

> answer4
 [1] B A B A A E D B A A B C B A E E A B B B
 
```{r compare}
table(answer1,answer2)
table(answer1,answer3)
table(answer2,answer3)
table(answer1,answer4)
table(answer2,answer4)
table(answer3,answer4)
```

answer 2 and 4 are identical.
- 1: 26 variables
- 2: random forest
- 3: 19 predictors only
- 4: full forest (160)
- 5: everything but NA

# Combine everything I got
```{r combine}
predDf <- data.frame(answer1,answer2,answer3,answer4,answer5,classe=testing$classe)
combModFit <- train(classe~.,data=predDf,method="glm")
combPred <- predict(combModFit,predDf)
```

## validation

pred1V <- predict(mod1,validation)
pred2V <- predict(mod2,validation)
predVDF <- data.frame(pred1=pred1V,pred2 =pred2V)
combPredV <- predict(combModFit,predVDF)

```{r validation}