---
title: "Machine Learning"
output: html_document
---

# Desicion trees
- find the node to split
- continue till pure

## impurity
- p = sum(all that vote)/whole set
- misclassification error: 1 - p
- 0 perfect purity
- 0.5 no purity

## gini index: 
- 1 - sum(pk^2)
- 0 perfect
- 0.5 none

## information gain
- -sum(p*log(p))
- 0 perfect
- 1 none


```{r decision tree}
data(iris)
library(ggplot2)
library(caret)
names(iris)
table(iris$Species)

inTrain <- createDataPartition(y=iris$Species,p=0.7,list=F)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
dim(training);dim(testing)
```

## petal width/sepal width
```{r petal sepal}
qplot(Petal.Width, Sepal.Width, colour=Species, data=training)
```

## plot it
```{r plot}
library(caret)
modFit <- train(Species ~ .,method="rpart",data=training)
print(modFit$finalModel)
plot(modFit$finalModel, uniform=TRUE,main="Classification Tree")
text(modFit$finalModel, use.n=TRUE, all=TRUE, cex=.8)
```

## pretty plot
```{r plot pretty}
library(rattle)
fancyRpartPlot(modFit$finalModel)
```

# Bagging
- bootstrap aggregating
- similar but litle bias
```{r bagging}
library(ElemStatLearn);data(ozone,package="ElemStatLearn")
ozone <- ozone[order(ozone$ozne),]
head(ozone)

# Bagged Loess
ll <- matrix(NA,nrow=10,ncol=155)
for(i in 1:10){
  ss <- sample(1:dim(ozone)[1],replace=T)
  ozone0 <- ozone[ss,]; ozone0 <- ozone0[order(ozone0$ozone),]
  loess0 <- loess(temperature ~ ozone, data=ozone0,span=0.2)
  ll[i,] <- predict(loess0,newdata=data.frame(ozone=1:155))
}
# loess is a smooth curve

# plot
plot(ozone$ozone, ozone$temperature,pch=19,cex=0.5)
for(i in 1:10){lines(1:155,ll[i,],col="grey",lwd=2)}
lines(1:155,apply(ll,2,mean),col="red",lwd=2)
```

## bagging in caret

- bagEarth
- treebag
- bagFDA
```{r bag caret}
predictors = data.frame(ozone=ozone$ozone)
temperature = ozone$temperature
treebag <- bag(predictors, temperature, B=10,
               bagControl = bagControl(fit = ctreeBag$fit,
                                       predict = ctreeBag$pred,
                                       aggregate = ctreeBag$aggregate))

plot(ozone$ozone, temperature, col="lightgrey",pch=19)
points(ozone$ozone, predict(treebag$fits[[1]]$fit,predictors),pch=19,col="red")
points(ozone$ozone, predict(treebag,predictors),pch=19,col="blue")
```

## parts of bagging
```{r bagging part}
ctreeBag$fit
ctreeBag$pred
ctreeBag$aggregate
```

# Random forest
- bootstrap samples
- bootstrap at each split
- grow multiple trees and vote
```{r random forest}
data(iris);library(ggplot2)
inTrain <- createDataPartition(y=iris$Species,p=0.7,list=F)
training <- iris[inTrain,]
testing <- iris[-inTrain,]

library(caret)
modFit <- train(Species ~ .,data=training,method="rf",prox=T)
modFit
getTree(modFit$finalModel,k=2)
```

## Class centers
```{r centers}
irisP <- classCenter(training[,c(3,4)],training$Species,modFit$finalModel$prox)
irisP <- as.data.frame(irisP)
irisP$Species <- rownames(irisP)
p <- qplot(Petal.Width,Petal.Length,col=Species,data=training)
p + geom_point(aes(x=Petal.Width,y=Petal.Length,col=Species),size=5,shape=4,data=irisP)
```

## Predict new values
```{r predict new}
pred <- predict(modFit,testing)
testing$predRight <- pred == testing$Species
table(pred, testing$Species)

# missed plots
qplot(Petal.Width, Petal.Length, colour=predRight,data=testing,main="newdata Predictions")
```

# Boosting
- weight and add up weak predictors
- take k classifiers
- sqn(sum(alpha*h))
- minimizer error on training set, updata weights
- adaboost on wikipedia

## Simple example
- weight strait lines
- gbm (trees)
- mboost(modeling)
- ada
```{r wage boost}
library(ISLR)
data(Wage)
library(ggplot2)
library(caret)
Wage <- subset(Wage,select=-c(logwage))
inTrain <- createDataPartition(y=Wage$wage,p=0.7,list=F)
training <- Wage[inTrain,]
testing <- Wage[-inTrain,]

modFit <- train(wage ~.,method="gbm",data=training,verbose=F)
print(modFit)
qplot(predict(modFit,testing),wage,data=testing)
```

# Model based prediction
- on bayes theorm
- pi is in advance
- f is a gaussian
- estimate mu,sigma from data
- classify to highet result of P

## Naive bayes
- assume all are independent
```{r iris bayes}
data(iris)
library(ggplot2)
names(iris)
table(iris$Species)
inTrain <- createDataPartition(y=iris$Species,p=0.7,list=F)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
dim(training); dim(testing)
```

## Build predictions
```{r bayes prediction}
modlda = train(Species ~.,data=training,method = "lda")
modnb = train(Species ~.,data=training,method="nb")
plda = predict(modlda,testing)
pnb = predict(modnb, testing)
table(plda, pnb)
```

## Comparison of results
```{r compare bayes}
equalPredictions = (plda==pnb)
qplot(Petal.Width, Sepal.Width,colour=equalPredictions,data=testing)
```

# Test

## Q1

1. Subset the data to a training set and testing set based on the Case variable in the data set. 
2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings. 
3. In the final model what would be the final model prediction for cases with the following variable values:
a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2 
b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100 
c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100 
d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2 
```{r q1}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)
library(dplyr)
#1 
segmentationOriginal <- as.data.frame(segmentationOriginal)
attach(segmentationOriginal)
training <- filter(segmentationOriginal, Case == "Train")
testing <- filter(segmentationOriginal, Case == "Test")
detach(segmentationOriginal)
dim(training);dim(testing)

#2
set.seed(125)
fitMod <- train(Class ~.,data=training,method="rpart")

#3
predict(fitMod,newdata = a)
library(rattle)
fancyRpartPlot(fitMod$finalModel)
fitMod$finalModel
```

a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2 
goes to PS
b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100 
goes to WS
c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100 
goes to PS
d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2 
can't decide

## Q2
large K means less bias and more variance.

## Q3
Load the olive oil data using the commands:
These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

newdata = as.data.frame(t(colMeans(olive)))
 
What is the resulting prediction? Is the resulting prediction strange? Why or why not?
0.005291005 0 0.994709 0 0 0 0 0 0. There is no reason why the result is strange.
0.005291005 0 0.994709 0 0 0 0 0 0. The result is strange because Area is a numeric variable and we should get the average within each leaf.
2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata
2.783. There is no reason why this result is strange.

```{r q3}
library(pgmm)
data(olive)
olive = olive[,-1]

modFit <- train(Area ~ .,method="rpart",data=olive)

newdata = as.data.frame(t(colMeans(olive)))

predict(modFit,newdata=newdata)
```


# Q4

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:

```{r q4}
library(ElemStatLearn)
library(caret)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]

set.seed(13234)
modFit <- train(chd ~ alcohol + obesity + tobacco + typea + ldl, data=trainSA,method="glm",family="binomial", )
attach(trainSA)
modFit <- train(chd ~ alcohol + obesity + tobacco + typea + ldl, data=trainSA,method = "glm",family="binomial")
detach(trainSA)

missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}

missClass(testSA$chd, predict(modFit,newdata=testSA))
missClass(trainSA$chd, predict(modFit,newdata=trainSA))
```

## Q5
```{r q5}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test) 
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
typeof(vowel.test$y)
str(vowel.train)

set.seed(33833)
fitMod <- train(y ~.,data=vowel.train,method="rf",prox=T)
?varimp
library(party)
varimp(predict(fitMod,newdata=vowel.test))
varimp(getTree(fitMod$finalModel,k=2))
varImp(fitMod)
```