---
title: "Statistics 101"
output: html_document
---

### Q1-1. Answer the below in 2 decimals.
- Population of Japan.
- Japan's annual budget
- GDP of Japan

<Answers>
- population: 130 million  
data from http://www.e-stat.go.jp/SG1/estat/List.do?lid=000001131732

- annual budget: 96 trillion yen  
data from http://www.bb.mof.go.jp/hdocs/bxss010bh27.html

- GDP: 524 trillion yen
data from http://www.esri.cao.go.jp/jp/sna/data/data_list/sokuhou/files/2014/qe144_2/pdf/jikei_1.pdf
```{r gdp}
setwd("/Users//shota/git//Coursera-Data-Science/Statistics/")
gdp <- read.csv("gaku-mg1442.csv")
```

### Q.2-2 calculate gini

```{r gini}
a <- c(0,3,3,5,5,5,5,7,7,10)
b <- c(0,1,2,3,5,5,7,8,9,10)
c <- c(3,4,4,5,5,5,5,6,6,7)
sum_avg <- function(x){
        sum = 0
        for (i in seq_along(x)){
                for (j in seq_along(x)){
                        sum = sum + abs(x[i] - x[j])
                }
        }
        sum/(length(x)^2)
}
sum_avg(a)
sum_avg(b)
sum_avg(c)
gini <- function(x){
        sum_avg(x)/(2 * mean(x))
}
mean(a)
mean(b)
mean(c)
gini(a)
gini(b)
gini(c)
```

### Q.3.1. Scatterplot the home ownership percentage and LDP party vote percentage

```{r vote}
vote <-  scan()
41.4 76.3 59.2 51.8 52.5 53.2 62.4 55.0 57.7 63.2 37.5 48.5 32.4
20.5 47.9 68.9 68.5 52.5 63.3 58.8 59.7 48.4 40.7 51.0 50.9 34.3
25.8 32.1 34.4 55.1 60.3 57.0 45.6 54.2 55.1 55.7 70.3 61.8 47.6
42.5 71.3 55.2 65.2 42.9 54.7 62.0 48.2 45.2


home <- scan()
52.8 71.2 72.6 63.7 81.3 81.8 70.9 74.0 73.2 72.9 66.7 65.7 43.7
55.5 79.6 85.7 75.3 80.5 73.0 77.0 77.5 69.2 60.0 78.2 79.5 61.8
49.6 59.6 72.1 71.0 76.3 72.8 71.8 60.7 67.0 71.8 71.2 68.3 68.5
54.8 76.0 65.8 69.4 66.9 69.7 71.2 59.6 62.4


pref <- c(
 "HOKKAIDO", "AOMORI", "IWATE", "MIYAGI", "AKITA", "YAMAGATA", "FUKUSHIMA",
 "IBARAKI", "TOCHIGI", "GUNMA", "SAITAMA", "CHIBA", "TOKYO", "KANAGAWA",
 "NIIGATA", "TOYAMA", "ISHIKAWA", "FUKUI", "YAMANASHI", "NAGANO", "GIHU",
 "SHIZUOKA", "AICHI", "MIE", "SHIGA", "KYOTO", "OSAKA", "HYOGO", "NARA",
 "WAKAYAMA", "TOTTORI", "SHIMANE", "OKAYAMA", "HIROSHIMA", "YAMAGUCHI",
 "TOKUSHIMA", "KAGAWA", "EHIME", "KOCHI", "FUKUOKA", "SAGA", "NAGASAKI",
 "KUMAMOTO", "OITA", "MIYAZAKI", "KAGOSHIMA", "OKINAWA", "ALL.japan") 

vote <- as.numeric(vote)
home <- as.numeric(home)
data <- data.frame(pref,vote,home)

library(ggplot2)
qplot(home, vote, data=data)
with(data,cor(home,vote))
```

#### Q.4.1.1 On rolling 4 dice, would you bet on there being or not being a 6?

```{r 4dice1six}
not_be <- (5/6)^4
be <- 1 - not_be
be > not_be
```

#### Q.4.1.2 Roll 2 dice 24 times, bet on would (6,6) come up or not
```{r 2dice24}
not_be <- (35/36)^24
be <- 1 - not_be
not_be > be
```

#### Q4.2 Roll 2 dice how many times till the probability that one will be 12 be more than 0.9?

```{r Huygens 11/14}
prob = 0
n = 0
while(prob < 0.9){
        n = n + 1
        prob = 1/36 * n  
}
print(n)
print(prob)
```

